### 2017-04-05 Project_Abstract.Rmd
Removed apostrophe

---
title: "A (sun)Shiny Recreation of the New York Times Weather Chart"
author: "Katie Will, Kiegan Rice, Amy Crawford"
output: pdf_document
---

### Abstract

In 1979 The New York Times published what would become a classic visualization of weather data that maps how temperatures compare to historic averages and records over the course of one year. This visualization has changed in several ways over the years, and in 2016 The New York Times published an online, interactive version of this chart, displaying 2015 temperature data for 3,116 world cities. In this project, we scrape historical data from [Weather Underground](https://www.wunderground.com/) for multiple cities for the years 1945 to 2017. With this data, we recreate the interactive chart published by The New York Times and allow for user interaction via a shiny app. Although we do not include all 3,116 cities, our app includes extra features that allow for users to view the data over the course of many years and choose the features they would like to include on their version of this iconic chart.