### 2017-04-08 Web_Scraping.Rmd
Add files via upload

---
title: 'Sun Shiny Weather: Web Scraping'
output: html_document
---

Reading in required packages:
```{r, include = F, warning = F, message = F}
library(tidyverse)
library(rvest)
library(lubridate)
```

Currently only for Des Moines airport, which starts 1945-01-01 with both 'actual' and 'average' temps.
Some airports don't have 'actual' temps until a later date.

Extracting temps for each days and record temps in one loop. (this is redundant) This will also take a couple hours to run. 
```{r, include = F, warning = F, message = F}
Date <- ymd("1945-01-01")  # set a starting date
day_temps <- list()        # empty list for min, max, and mean temps, actual and historical averages
record_temps <- list()     # empty list for record temps
                           # one item in the list for each day
i <- 1                     # counter for the list item

# Need to create the url which changes for each day (and each airport)
url_begin <- "https://www.wunderground.com/history/airport/" 
airport <- "KDSM/"                  # This will change in loop (eventually)
url_end <- "/DailyHistory.html"

  while(Date < Sys.Date()){
    date <- gsub("-", "/", Date)      
    url <- paste(url_begin, airport, date, url_end, sep = "") # Pasting together parts of url
    html <- read_html(url)

    # grabbing the "Actual" daily temperatures (mean, min, and max)
    actual_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(2) , #historyTable tr:nth-child(3) td:nth-child(2), #historyTable tr:nth-child(2) td:nth-child(2)") %>% html_text %>% 
      gsub("-", "NA", .) %>% gsub("°F", "", .) %>% gsub("\n", "", .) %>% gsub("[[:blank:]]", "", .)

    # grabbing the historical "Average" temperatures (mean, min, and max)
    average_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(3) , #historyTable tr:nth-child(3) :nth-child(3), #historyTable tr:nth-child(2) td:nth-child(3)") %>% html_text %>% 
      gsub("-", "NA", .) %>% gsub("°F", "", .) %>% gsub("\n", "", .) %>% gsub("[[:blank:]]", "", .)

    # grabbing the labels for the temperature values
    temp <- html %>% 
      html_nodes("tr:nth-child(4) .indent , tr:nth-child(3) .indent, tr:nth-child(2) .indent") %>% 
      html_text()

    # grabbing the record min and max for each day
    # this is redundant and may need to be fixed for efficiency purposes
    record <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(4) , #historyTable tr:nth-child(3) td:nth-child(4)") %>%
      html_text()
    record_vals <- parse_number(record)
    record_yrs <- record %>% gsub("^[^\\(]+", "", .) %>% parse_number(.)

    # putting data into a data frame which then makes up one item in the master list
    # two different lists (one for daily temps and one for the record temps)
    day_temps[[i]] <- data.frame(day = Date,
                          mean_max_min = rep(temp, times = 2),
                          actual_ave = rep(c("Actual", "Historical Average"), each = 3),
                          value = as.numeric(c(actual_vals, average_vals)))

    record_temps[[i]] <- data.frame(day = "08-11",
                               year = record_yrs,
                               record_type = c("Max", "Min"),
                               value = record_vals)

    Date <- Date+1
    i <- i+1
  }

# binding together all items of the list
DSM_temps <- dplyr::bind_rows(day_temps)
DSM_records <- dplyr::bind_rows(record_temps)
```

To fix the redundancy, split the loop into two.

This loop extracts the record temps (so a min and max for 366 days)
```{r}
# Two while loops to fix the redundant records

# loop to grab record min and max from the web (note this is only for 366 days)
Date <- ymd("2016-01-01")
record_temps <- list() 
i <- 1

# Need to create the url which changes for each day (and each airport)
url_begin <- "https://www.wunderground.com/history/airport/" 
airport <- "KDSM/"                  # This will change in loop (eventually)
url_end <- "/DailyHistory.html"

  while(Date < ymd("2017-01-01")){
    date <- gsub("-", "/", Date)      
    url <- paste(url_begin, airport, date, url_end, sep = "") # Pasting together parts of url
    html <- read_html(url)

    # grabbing the record min and max for each day
    # this is redundant and may need to be fixed for efficiency purposes
    record <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(4) , #historyTable tr:nth-child(3) td:nth-child(4)") %>%
      html_text()
    record_vals <- parse_number(record)
    record_yrs <- record %>% gsub("^[^\\(]+", "", .) %>% parse_number(.)

    record_temps[[i]] <- data.frame(month_number = month(Date, label = F),
                                    month_name = month(Date, label = T),
                                    day = day(Date),
                                    record_year = record_yrs,
                                    record_type = c("Max", "Min"),
                                    value = record_vals)

    Date <- Date+1
    i <- i+1
  }

DSM_records <- dplyr::bind_rows(record_temps)

```

Plots of record temps
```{r}
DSM_records <- DSM_records %>% mutate(day_of_year = 1:nrow(DSM_records), date = as.Date(paste(month_name, day, sep = "-"), format = "%b-%d"))

plot1 <-  DSM_records %>% ggplot(aes(y = value, x = day_of_year, colour = record_type)) +
                            geom_point(aes(text = record_year))
plotly::ggplotly(plot1)

plot1 <-  DSM_records %>% ggplot(aes(y = value, x = month_number, colour = record_type)) + 
                            geom_point(aes(text = record_year)) +
                            scale_x_discrete(name = "Month", 
                                             breaks = data$month_number,
                                             labels = data$month_name
                                             )

```

This loop extracts the temps for each day (so a min, max, mean for 'actual' and 'average' for approx. 26,280 days)
```{r}
# loop to grab daily temps from the web (note this is approximately 26,000 days for DSM)
Date <- ymd("1945-01-01")  # set a starting date
day_temps <- list()        # empty list for min, max, and mean temps, actual and historical averages
i <- 1                     # counter for the list item

# Need to create the url which changes for each day (and each airport)
url_begin <- "https://www.wunderground.com/history/airport/" 
airport <- "KDSM/"                  # This will change in loop (eventually)
url_end <- "/DailyHistory.html"

  while(Date < Sys.Date()){
    date <- gsub("-", "/", Date)      
    url <- paste(url_begin, airport, date, url_end, sep = "") # Pasting together parts of url
    html <- read_html(url)

    # grabbing the "Actual" daily temperatures (mean, min, and max)
    actual_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(2) , #historyTable tr:nth-child(3) td:nth-child(2), #historyTable tr:nth-child(2) td:nth-child(2)") %>% html_text %>% 
      gsub("-", "NA", .) %>% gsub("°F", "", .) %>% gsub("\n", "", .) %>% gsub("[[:blank:]]", "", .)

    # grabbing the historical "Average" temperatures (mean, min, and max)
    average_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(3) , #historyTable tr:nth-child(3) :nth-child(3), #historyTable tr:nth-child(2) td:nth-child(3)") %>% html_text %>% 
      gsub("-", "NA", .) %>% gsub("°F", "", .) %>% gsub("\n", "", .) %>% gsub("[[:blank:]]", "", .)

    # grabbing the labels for the temperature values
    temp <- html %>% 
      html_nodes("tr:nth-child(4) .indent , tr:nth-child(3) .indent, tr:nth-child(2) .indent") %>% 
      html_text()

    # putting data into a data frame which then makes up one item in the master list
    # two different lists (one for daily temps and one for the record temps)
    day_temps[[i]] <- data.frame(year = year(Date), 
                                 month = month(Date, label = T),
                                 day = day(Date),
                                 mean_max_min = rep(temp, times = 2),
                                 actual_ave = rep(c("Actual", "Historical Average"), each = 3),
                                 value = as.numeric(c(actual_vals, average_vals)))

    Date <- Date+1
    i <- i+1
  }

DSM_temps <- dplyr::bind_rows(day_temps)

```

*Attempt to optimize:* 
```{r}
library(reshape)
# Attempt at better efficiency... using html_table

Date <- ymd("1945-01-01")

# URL code stays the same
    url_begin <- "https://www.wunderground.com/history/airport/" 
    airport <- "KDSM/"                  # This will change in loop (eventually)
    date <- gsub("-", "/", Date)      
    url_end <- "/DailyHistory.html"
    url <- paste(url_begin, airport, date, url_end, sep = "") # Pasting together parts of url
    html <- read_html(url)
    tables <- html %>% html_table(fill = TRUE)

    sm_table <- tables[[1]][2:4,1:3]
    sm_table <- setNames(sm_table, c("Temperature", "Actual", "Average"))
    rec_table <- tables[[1]][3:4, c(1,4)] 
    rec_table <- rec_table %>% setNames(., c("Temperature", "Record")) %>%
                      transform(., Record = colsplit(Record, split = "\\\n", names = c('temp', 'year')))
    sm_table
    rec_table

```