### 2017-04-08 Web_Scraping_attempt2.Rmd
Add files via upload

---
title: 'Web Scraping: attempt 2'
author: "Katie Will"
date: "April 8, 2017"
output: html_document
---

*html_session within function *

4/8/17  1:44pm    SYSTEM.TIME:  336.107 seconds elapsed

```{r}
#url_begin <- "https://www.wunderground.com/history/airport/KDSM/" 
link_dates <- gsub(pattern = "-", replacement = "/", x = seq(from = as.Date("1945-01-01"), to = Sys.Date(), by = "days"))
links <- paste0(link_dates, "/DailyHistory.html?req_city=&req_state=&req_statename=&reqdb.zip=&reqdb.magic=&reqdb.wmo=")

#date <- "2017/04/08"
#links <- paste0(date, "/DailyHistory.html?req_city=&req_state=&req_statename=&reqdb.zip=&reqdb.magic=&reqdb.wmo=")

#s <- html_session("https://www.wunderground.com/history/airport/")
#html_form(s)

grab_temps <- function(link){
  s <- html_session("https://www.wunderground.com/history/airport/")
  ss <- s %>% jump_to(paste0("KDSM/",link))
  html <- read_html(ss)

  actual_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(2) , #historyTable tr:nth-child(3) td:nth-child(2), #historyTable tr:nth-child(2) td:nth-child(2)") %>% html_text() %>% gsub("\n", "", .) %>% parse_number(., na = c("-", "NA"))

    # grabbing the historical "Average" temperatures (mean, min, and max)
  average_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(3) , #historyTable tr:nth-child(3) :nth-child(3), #historyTable tr:nth-child(2) td:nth-child(3)") %>% html_text() %>% gsub("\n", "", .) %>% parse_number(., na = c("-", "NA"))

    # grabbing the labels for the temperature values
    temp <- html %>% 
      html_nodes("tr:nth-child(4) .indent , tr:nth-child(3) .indent, tr:nth-child(2) .indent") %>% 
      html_text()

    return(data.frame(day = as.Date(substring(text = link, first = 1, last = 10)),
                      mean_max_min = as.character(rep(temp, times = 2)),
                      actual_ave = as.character(rep(c("Actual", "Historical Average"), each = 3)),
                      value = as.numeric(c(actual_vals, average_vals))))
}

## WORKING ###
df2 <- data.frame(Links = links[1:365])
system.time(df2 %>% nest() %>% mutate(daily_temps = Links %>% purrr::map(grab_temps))) #%>% select(-Links) %>% unnest()
##############

```

*Pasting togeter url and then using read_html*

4/8/17  1:52pm   SYSTEM.TIME:   343.802 elapsed

```{r}

grab_temps2 <- function(date){
    url <- paste0("https://www.wunderground.com/history/airport/", "KDSM/", date, "/DailyHistory.html") # Pasting together parts of url
    html <- read_html(url)

  actual_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(2) , #historyTable tr:nth-child(3) td:nth-child(2), #historyTable tr:nth-child(2) td:nth-child(2)") %>% html_text() %>% gsub("\n", "", .) %>% parse_number(., na = c("-", "NA"))

    # grabbing the historical "Average" temperatures (mean, min, and max)
  average_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(3) , #historyTable tr:nth-child(3) :nth-child(3), #historyTable tr:nth-child(2) td:nth-child(3)") %>% html_text() %>% gsub("\n", "", .) %>% parse_number(., na = c("-", "NA"))

    # grabbing the labels for the temperature values
    temp <- html %>% 
      html_nodes("tr:nth-child(4) .indent , tr:nth-child(3) .indent, tr:nth-child(2) .indent") %>% 
      html_text()

    return(data.frame(day = as.Date(substring(text = link, first = 1, last = 10)),
                      mean_max_min = as.character(rep(temp, times = 2)),
                      actual_ave = as.character(rep(c("Actual", "Historical Average"), each = 3)),
                      value = as.numeric(c(actual_vals, average_vals))))
}

link_dates <- gsub(pattern = "-", replacement = "/", x = seq(from = as.Date("1945-01-01"), to = Sys.Date(), by = "days"))
airport <- "KDSM/"

## WORKING ###
df3 <- data.frame(Dates = link_dates[1:365])
system.time(df3 %>% nest() %>% mutate(daily_temps = Dates %>% purrr::map(grab_temps2))) #%>% select(-Links) %>% unnest()
##############

```