### 2017-04-09 static visual.Rmd
create static visual file

---
title: "static visual"
author: "Kiegan Rice"
date: "April 9, 2017"
output: html_document
---

*install stuff*
```{r}
# installing ish
library(tidyverse)
library(beepr)
```

*read data*
```{r}
temps <- read.csv(file = "./raw-data/DSM_temps.csv")
head(temps)
```

*get dates in right format*
```{r}
dates_2016 <- gsub(pattern = "-", replacement = "/", x = seq(from = as.Date("2016-01-01"), 
                                                                  to = as.Date("2016-12-31"), by = "days"))

temps <- temps %>% mutate(date_2016 = lubridate::ymd(paste("2016", month, day, sep = "-")))
```

*base of plot*
```{r}
base <- temps %>% ggplot(aes(x = date_2016))
base
```

*record layer*
```{r}
data_record <- temps %>% filter(actual_ave_rec == "Record") %>% group_by(date_2016) %>%
  spread(key = mean_max_min, value = value)

layer_record <- geom_point(data = data_record, aes(x = date_2016+1, y = data_record$`Min Temperature`))
layer_record <- geom_rect(data = data_record, aes(xmin = date_2016-.5, xmax = date_2016+.5, ymin = data_record$`Min Temperature`, ymax = data_record$`Max Temperature`, fill = "black"))

base + layer_record
```