### 2017-04-10 static visual.Rmd
static fix

---
title: "static visual"
author: "Kiegan Rice"
date: "April 9, 2017"
output: html_document
---

*install stuff*
```{r}
# installing ish
library(tidyverse)
library(beepr)
```

*read data*
```{r}
temps <- read.csv(file = "./raw-data/DSM_temps.csv")
head(temps)
```

*get dates in right format*
```{r}
dates_2016 <- gsub(pattern = "-", replacement = "/", x = seq(from = as.Date("2016-01-01"), 
                                                                  to = as.Date("2016-12-31"), by = "days"))

temps <- temps %>% select(-X) %>% mutate(date_2016 = lubridate::ymd(paste("2016", month, day, sep = "-")))
```

*base of plot*
```{r}
base <- temps %>% ggplot(aes(x = date_2016)) + 
  theme_classic() +
  theme(panel.grid.major.x = element_line(colour = c("black", rep("grey", times = 11)), linetype = c(0, rep(2, times = 11))),
        axis.text.x = element_text(hjust = -0.9), 
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_line(colour = "darkgrey")) +
  scale_x_date(date_breaks = "1 month", 
               date_labels = "%b", 
               expand=c(0,0)) + 
  scale_y_continuous(breaks = seq(-20, 120, by = 20), limits= c(-40, 120)) + 
  xlab("") +
  ylab("")

text_layer_description <- annotate("text", 
                                   label = "Bars represent the range between the daily high and low.",
                                   x = lubridate::ymd("2016-07-24"),
                                   y = -28, hjust = "center", size = 2.3)

base <- base + text_layer_description
base 
```

*record layer*
```{r, fig.width = 10}
data_record <- temps %>% select(-date, -year) %>% 
  filter(actual_ave_rec == "Record", mean_max_min != "Mean Temperature") %>% 
  group_by(month, day) %>% spread(key = mean_max_min, value = value)

layer_record <- geom_rect(data = data_record, aes(xmin = date_2016, xmax = date_2016+1, ymin = data_record$`Min Temperature`, ymax = data_record$`Max Temperature`), fill = "#E6E5DD", alpha = 0.8)

legend_layer_record <- geom_rect(aes(xmin = lubridate::ymd("2016-07-22"), 
                                     xmax =  lubridate::ymd("2016-07-27"), 
                                     ymin = -15, 
                                     ymax = 15), 
                                 fill = "#E6E5DD", alpha = 0.8)
text_layer_record <-  annotate("text",
                               label = c("Record Low", "Record High"), 
                               x =  c(lubridate::ymd("2016-07-22"), lubridate::ymd("2016-07-22")), 
                               y = c(-15, 15), 
                               hjust = "right", 
                               size = 2.3)

records_plot <- base + 
  layer_record + legend_layer_record + text_layer_record

records_plot
```

*historical averages layer*
```{r}
data_average <- temps %>% select(-date, -year) %>% 
  filter(actual_ave_rec == "Historical Average", mean_max_min != "Mean Temperature") %>% 
  group_by(month, day) %>% spread(key = mean_max_min, value = value)

layer_average <- geom_rect(data = data_average, aes(xmin = date_2016, xmax = date_2016+1, 
                                                   ymin = data_average$`Min Temperature`, ymax = data_average$`Max Temperature`), 
                           fill = "#BCACAC", alpha = 0.8)

legend_layer_average <- geom_rect(aes(xmin = lubridate::ymd("2016-07-24")-0.5, 
                                     xmax =  lubridate::ymd("2016-07-25")+0.5, 
                                     ymin = -10, 
                                     ymax = 10), 
                                 fill = "#BCACAC", alpha = 0.8)

text_layer_average <-  annotate("text",
                                label = "Normal range", 
                                x =  lubridate::ymd("2016-07-19"), 
                                y = 0, 
                                hjust = "right", 
                                size = 2.3)

line_layer_average_ver <- geom_segment(aes(x = lubridate::ymd("2016-07-20"), 
                                       xend = lubridate::ymd("2016-07-20"), 
                                       y = -10, 
                                       yend = 10), 
                                   size = 0.01)

line_layer_average_hor1 <- geom_segment(aes(x = lubridate::ymd("2016-07-20"),
                   xend = lubridate::ymd("2016-07-22"),
                   y = 10,
                   yend = 10),
               size = 0.01)

line_layer_average_hor2 <- geom_segment(aes(x = lubridate::ymd("2016-07-20"),
                   xend = lubridate::ymd("2016-07-22"),
                   y = -10,
                   yend = -10),
               size = 0.01)

averages_plot <- base + layer_average + 
  legend_layer_average + text_layer_average + 
  line_layer_average_ver + line_layer_average_hor1 + line_layer_average_hor2

records_averages_plot <- records_plot + layer_average + 
  legend_layer_average + text_layer_average + 
  line_layer_average_ver + line_layer_average_hor1 + line_layer_average_hor2

averages_plot
records_averages_plot
```

*actual temps layer*
```{r, fig.width = 12, fig.height=4}
data_actual <- temps %>% 
  filter(actual_ave_rec == "Actual", year == 2015) %>% 
  group_by(month, day) %>% spread(key = mean_max_min, value = value)

layer_actual <- geom_rect(data = data_actual, aes(xmin = date_2016-.5, xmax = date_2016+.5, 
                                                   ymin = data_actual$`Min Temperature`, ymax = data_actual$`Max Temperature`), 
                           fill = "maroon")
layer_white_ylines <- geom_hline(yintercept = seq(-20, 100, by = 20), colour = "white", lwd = 0.1)

legend_layer_actual <- geom_rect(aes(xmin = lubridate::ymd("2016-07-24")-0.05, 
                                     xmax =  lubridate::ymd("2016-07-25")+0.15, 
                                     ymin = -5, 
                                     ymax = 12), 
                                 fill = "maroon")

text_layer_actual <-  annotate("text",
                               label = c("Actual Low", "Actual High"), 
                               x =  c(lubridate::ymd("2016-07-29"), lubridate::ymd("2016-07-29")), 
                               y = c(-5, 12), 
                               hjust = "left", 
                               size = 2.3)

line_layer_actual_hor1 <- geom_segment(aes(x = lubridate::ymd("2016-07-24"),
                   xend = lubridate::ymd("2016-07-28"),
                   y = -5,
                   yend = -5),
               size = 0.01)

line_layer_actual_hor2 <- geom_segment(aes(x = lubridate::ymd("2016-07-24"),
                   xend = lubridate::ymd("2016-07-28"),
                   y = 12,
                   yend = 12),
               size = 0.01)

actual_plot <- base + layer_actual + legend_layer_actual + text_layer_actual + line_layer_actual_hor1 + line_layer_actual_hor2

all_plot <- records_plot + 
  layer_average + 
  legend_layer_average + text_layer_average + 
  line_layer_average_ver + line_layer_average_hor1 + line_layer_average_hor2 + 
  layer_white_ylines +
  layer_actual + legend_layer_actual + text_layer_actual + line_layer_actual_hor1 + line_layer_actual_hor2

records_actual_plot <- records_plot + 
    layer_white_ylines +
  layer_actual + legend_layer_actual + text_layer_actual + line_layer_actual_hor1 + line_layer_actual_hor2

averages_actual_plot <- averages_plot + 
    layer_white_ylines +
  layer_actual + legend_layer_actual + text_layer_actual + line_layer_actual_hor1 + line_layer_actual_hor2

records_plot
averages_plot
actual_plot
records_averages_plot
records_actual_plot
averages_actual_plot
all_plot
beep(8)
```