### 2017-04-18 Web_Scraping_Final.Rmd

---
title: 'Web Scraping: attempt 2'
author: "Katie Will"
date: "April 8, 2017"
output: html_document
---
### Notes for use:  
\begin{enumerate}
\item Change "airport" (line 46, line 90)
\item Change name of data frame when calling function to e.g. (DSM, IAH, FAR)
\item Make sure those name changes are in each place it gets called
\item Change `city = "Des Moines, IA"` to correct city, state in mutate function
\item Update save function to have the correct file and .csv name
\end{enumerate}

*daily temperatures (approximately 26,000 days for DSM)*

```{r}
library(beepr)
library(tidyverse)
library(rvest)
dates_all <- gsub(pattern = "-", replacement = "/", x = seq(from = as.Date("1948-01-01"), 
                                                                 to = as.Date("2017-03-31"), by = "days"))
links_all <- paste0(dates_all, "/DailyHistory.html")

grab_act_temps <- function(link, airport){
  s <- html_session("https://www.wunderground.com/history/airport/")
  ss <- s %>% jump_to(paste0(airport, "/", link))
  html <- read_html(ss)

   actual_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(2) , #historyTable tr:nth-child(3) td:nth-child(2), #historyTable tr:nth-child(2) td:nth-child(2)") %>% html_text() %>% gsub("\n", "", .) %>% parse_number(., na = c("-", "NA"))

    # grabbing the labels for the temperature values
    temp <- html %>% 
      html_nodes("tr:nth-child(4) .indent , tr:nth-child(3) .indent, tr:nth-child(2) .indent") %>% 
      html_text()

    return(data.frame(date = as.Date(substring(text = link, first = 1, last = 10)),
                      mean_max_min = as.character(temp),
                      actual_ave_rec = as.character(rep("Actual", times = 3)),
                      value = as.numeric(actual_vals)))
}

airport <- "KDET"
DET_daily <- data.frame(Links = links_all)  
DET_daily <- DET_daily %>% nest() %>% 
  mutate(daily_temps = Links %>% purrr::map2(., airport, grab_act_temps)) %>%
                  select(-Links) %>% unnest(data) %>% unnest(daily_temps) %>% as.data.frame()
beepr::beep(8)
```

*average and record temperatures (366 days)*
```{r}

dates_2016 <- gsub(pattern = "-", replacement = "/", x = seq(from = as.Date("2016-01-01"), 
                                                                  to = as.Date("2016-12-31"), by = "days"))
links_2016 <- paste0(dates_2016, "/DailyHistory.html")

grab_ave_rec_temps <- function(link, airport){
  s <- html_session("https://www.wunderground.com/history/airport/")
  ss <- s %>% jump_to(paste0(airport, "/", link))
  html <- read_html(ss)

    # grabbing the historical "Average" temperatures (mean, min, and max)
    average_vals <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(3) , #historyTable tr:nth-child(3) :nth-child(3), #historyTable tr:nth-child(2) td:nth-child(3)") %>% html_text() %>% gsub("\n", "", .) %>% parse_number(., na = c("-", "NA"))

    # grabbing the labels for the temperature values
    temp <- html %>% 
      html_nodes("tr:nth-child(4) .indent , tr:nth-child(3) .indent, tr:nth-child(2) .indent") %>% 
      html_text()

    record <- html %>% 
      html_nodes("#historyTable tr:nth-child(4) td:nth-child(4) , #historyTable tr:nth-child(3) td:nth-child(4)") %>%
      html_text()
    record_vals <- c("NA", parse_number(record))
    record_yrs <- record %>% gsub("^[^\\(]+", "", .) %>% parse_number()

    return(data.frame(date = as.Date(c(rep(substring(text = link, first = 1, last = 10), times = 3), "NA",
                              paste0(record_yrs, substring(text = link, first = 5, last = 10)))),
                      mean_max_min = as.character(rep(temp, times = 2)),
                      actual_ave_rec = as.character(rep(c("Historical Average", "Record"), each = 3)),
                      value = as.numeric(c(average_vals, record_vals))))
}

airport <- "KDET"
DET_ave_rec <- data.frame(Links = links_2016)  
DET_ave_rec <- DET_ave_rec %>% nest() %>% 
  mutate(ave_rec_temps = Links %>% purrr::map2(., airport, grab_ave_rec_temps)) %>% 
  select(-Links) %>% unnest(data) %>% unnest(ave_rec_temps) %>% as.data.frame()

```

*put two datasets together and add some elements*
```{r}
head(DET_daily)
#names(DSM_daily) <- c("date", "mean_max_min", "actual_ave_rec", "value") # just one fix, already fixed in other code
head(DET_ave_rec)
#names(DSM_ave_rec) <- c("date", "mean_max_min", "actual_ave_rec", "value") # just one fix, already fixed in other code
DET_temps <- rbind(DET_daily, DET_ave_rec)

DET_temps <- DET_temps %>% mutate(airport = airport)
DET_temps <- DET_temps %>% mutate(city = "Detroit, MI") # change this each time

DET_temps <- DET_temps %>% mutate(day = lubridate::day(date),
                                  month = lubridate::month(date),
                                  year = lubridate::year(date))
```

*save final dataset*
```{r}
write.csv(DET_temps, file = "raw-data/DET_temps.csv")
```